---
title: "homicides-Col-municipalitites-Exploration"
output: 
  html_notebook:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
    number_sections: true
    code_folding: "hide"
    theme: "cosmo"
    highlight: "monochrome"
---


```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(summarytools)
library(xtable)
library(knitr)
library(ExPanDaR)
library(plotly)
library(REAT)
library(DT)

# Change the presentation of decimal numbers to 4 and avoid scientific notation
options(prompt="R> ", digits=2, scipen=999)
```



# Import the data

Via the `tidyverse` package

 using the code because the name of the municipalities is in spanish "locale = locale(encoding = "ISO-8859-1") "
 
 
```{r}
dat <- read_csv("homicides_Col_municipalities_2010_18.csv",    locale = locale(encoding = "ISO-8859-1"))
dat
View(dat)

```

Use the package `DT` to present the data frame in an interactive way

```{r}
dat %>% 
  datatable(filter = "top") %>%
  formatRound(colnames(dat), 2) 
```



# Decribe data

Use the package `summarytools` to describe the data

```{r}
print(dfSummary(dat), method = "render")
```

## Identify missing values


```{r}
prepare_missing_values_graph(dat_hr, ts_id = "year")
```

# Explore data

Use the package `ExPanDaR` to explore panel data. However, to use it, we first need to reshape the dataset and make it long form. 

```{r}
dat_hr <- dat %>%
  select(DPMP, starts_with("hr")) %>% 
  gather(hryear, hrate, hrate10:hrate18) %>% 
  mutate(year= 2000+ as.integer(substr(hryear,6,7))) %>%
  select(1,4,3)

dat_hr
```


## Plot: Dispersion dynamics


```{r}


graph3 <- prepare_quantile_trend_graph(dat_hr, ts_id = "year", quantiles = c(0.01, 0.05, 0.1,0.2, 0.25, 0.5, 0.75, 0.95), var = "hrate")
graph3$plot

```

## Plot: overall trend

```{r}
graph1 <- prepare_trend_graph(dat_hr, ts_id = "year", "hrate")
graph1$plot
```



